\LaTeX\ Author Guidelines for ICCV Proceedings

### 2022-05-19 chapters/supplementary.tex
Update to CVPR2022 style
Camera ready changes

@@

\paragraph*{Update rules} During the agglomerative process, the interaction between adjacent clusters has to be properly updated and recomputed, as shown in Algorithm \ref{main_alg}.  
An efficient way of implementing these updates can be achieved by representing the agglomeration as a sequence of \emph{edge contractions} in the graph. Given a graph $\mathcal{G}(V,E,\cost)$ and a clustering $\Pi$, we define the associated \emph{contracted graph} $\tilde{\mathcal{G}}_\Pi(\tilde{V}, \tilde{E}, \tilde{\cost})$, such that there exists exactly one representative node $|\tilde{V} \cap S| = 1$ for every cluster $S \in \Pi$ . Edges in $\tilde{E}$ represent adjacency-relationships between clusters 
and the signed edge weights $\tilde{\cost}_e$ are given by inter-cluster interactions $\tilde{\cost}(e_{uv})=\interact_{S_u,S_v}$, where $S_u$ denotes the clustering including node $u$. 
For the linkage criteria tested in this article, when two clusters $S_u$ and $S_v$ are merged, the interactions between the new cluster $S_u \cup S_v$ and each of its neighbors depend only on the previous interactions involving $S_u$ and $S_v$. Thus, we can recompute these interactions by using an \emph{update rule} $f$ that does not involve any loop over the edges of the original graph $\mathcal{G}$:
\begin{align}
  \interact(S_u \cup S_v,S_t) =& f\Big[ \interact(S_u,S_t), \interact(S_v,S_t) \Big] \\
  =& f(\tilde{\cost}(e_{ut}), \tilde{\cost}(e_{vt})) 
\end{align}
In Fig. \ref{fig:edge_contraction_and_contr_graph} we show an example of edge contraction and in Table \ref{tab:linkage_criteria_explicit} we list the update rules associated to the linkage criteria we introduced in Table \ref{tab:linkage-criteria}.

@@

@@

\begin{lemma} \label{lemma:absMax_and_complete_not_positive}
If \algname{} Algorithm \ref{main_alg} with \textbf{AbsMax linkage criteria} enforces a constraint between two clusters in \emph{Phase 1}, then the interaction between the clusters will never become positive over the course of the following agglomeration steps.
\end{lemma}
\begin{proof}
During the agglomeration the interaction between two clusters can only increase in absolute value. Thus, the negative interaction $\interact{}(S_i,S_j)<0$ between two constrained clusters can possibly become positive over the course of next agglomeration steps only if there is at least another pair of clusters in the graph that has a positive interaction $\interact{}(S_l,S_t)>0$ higher in absolute value: $|\interact{}(S_l,S_t)| > |\interact{}(S_i,S_j)|$.
If such clusters $S_l, S_t$ with positive interaction exist, we note that they must also be constrained (in the opposite case, the algorithm would have already merged them before to constrain $S_i$ and $S_j$, because their priority is higher). In other words, a constrained negative interaction can become positive only if there is already another positive constrained interaction: but this can never be the case because initially all constrained interactions are negative.
\end{proof}

@@

The constrained versions of \algname{} (HCC-Avg and HCC-Single) are also not weight-shift invariant: here, the algorithm merges or constrains clusters in a given order, depending on the absolute interactions $|\interact{}(S_i, S_j)|$ between clusters; so, when edge weights are shifted by a constant $\alpha$, the sorting by absolute value can change arbitrarily together with the agglomeration order, as we show in the counter-example of Fig.~\ref{fig:counter_examples_shift_inv}.
Similarly, the Mutex Watershed algorithm is not weight-shift invariant because it uses a linkage criterion that compares weights by their absolute values (see again counter-example in Fig.~\ref{fig:counter_examples_shift_inv}.
\end{proof}

@@

\algname{} puts clusters in a priority queue (Algorithm \ref{main_alg}, lines 5 and 15) and merges them starting from those with the highest interaction (lines 9, 19, and 26). However, the priority queue is updated each time two clusters are merged (lines 10, 20, and 27). Thus, to ensure a monotonously decreasing merging order, updated interactions involving a merged cluster should always be lower or equal than previously existing interactions (\textbf{condition 1}):
\begin{align}\label{eq:condition1}
& \forall S_i \in \Pi \setminus \{S_1, S_2\}, \nonumber \\
\interact{}(S_1 \cup S_2, S_i) &\leq \max \{ \interact{}(S_1, S_i), \interact{}(S_2, S_i)\} 
\end{align}
where $\Pi$ is a clustering, $\interact{}$ is a linkage criteria, and $S_1,S_2\in \Pi$ are two clusters merged by the algorithm at a given iteration. If this condition is true then, in the following iterations, \algname{} can only merge clusters with lower (or equal) interaction values.

@@

\subsection{Mutex Watershed on SSBM graphs}
\begin{prop} \label{prop:MWS_on_SSBM}
Consider a graph generated by an Erd\H os-R\'enyi signed stochastic block model (SSBM) as described in Section \ref{sec:clustering_problems}, with $N$ nodes, edges added with probability $p$, sign-flip probability $\eta<0.5$, $k$ ground-truth clusters, and edge weights Gaussian-distributed with standard deviation $\sigma$. Then, at every iteration, \algname{} with Absolute Maximum linkage (or, in other words, the Mutex Watershed algorithm) always makes a mistake with at least probability $\eta$. 
\end{prop}
\begin{proof}
Thanks to Lemma \ref{lemma:absMax_and_complete_property} we know that \algname{} with Absolute Maximum linkage returns the same clustering whether or not cannot-link-constraints are used. Thus, in the following, we prove the proposition considering the version enforcing constraints.
Let us consider a generic iteration of the algorithm, where two clusters $S_\alpha$ and $S_\beta$ have the highest priority and are popped from priority queue. Then, the MWS algorithm will either merge or constrain them depending on the fact that their interaction $\interact_{\mathrm{AbsMax}}(S_\alpha,S_\beta)$ is  positive or negative (note that, with AbsMax linkage, an interaction can never be positive and constrained, as shown in Lemma \ref{lemma:absMax_and_complete_property}).
By construction of the SSBM, every edge $e\in E$ in the graph has a absolute weight distributed as $|w_e|\sim \mathcal{N}(1,\sigma^2)$. Thus, every edge $e'\in(S_{\alpha} \times S_{\beta})\cap E$ connecting the two clusters has the same probability to have the highest absolute weight, and the sign of the interaction $\interact_{\mathrm{AbsMax}}(S_\alpha,S_\beta)$ will only depend on the sign of this highest edge. Therefore, the probability that the MWS merges two clusters is simply given by the fraction of positive weighted edges connecting them.

@@

So far we have been talking about attractive and repulsive connections according to the ground truth labels. In our SSBM however every edge has a uniform probability $\eta$ to have its sign flip, so  the actual number of attractive interactions connecting the two clusters will be instead given by the sum of the true attractive interactions $\nBE^+_{\mathrm{nf}}\sim \mathcal{B}(\nBE^+,1-\eta)$ that have not been flipped, plus the true negative interactions $\nBE^-_{\mathrm{f}}\sim \mathcal{B}(\nBE^-,\eta)$ that have been flipped.
Putting everything together, given two clusters with $\nBE^+$ true attractive interactions and $\nBE^-$ true negative ones, the highest-absolute-weight edge connecting them has the following probability to be positive:
\begin{align}\label{eq:mws_prob}
 \mathbb{P}[\interact_{\mathrm{AbsMax}}&(S_\alpha,S_\beta)>0;\nBE^+,\nBE^-] = \nonumber \\
=&\sum_{\nBE^+_{\mathrm{nf}}=0}^{\nBE^+}\sum_{\nBE^-_{\mathrm{f}}=0}^{\nBE^-} 
\mathcal{B}(\nBE^-_{\mathrm{f}};\nBE^-,\eta) \mathcal{B}(\nBE^+_{\mathrm{nf}};\nBE^+,1-\eta)\cdot  \nonumber\\
& \qquad \qquad  \cdot \left(\frac{\nBE^+_{\mathrm{nf}}+\nBE^-_{\mathrm{f}}}{\nBE^+ + \nBE^-} \right) \nonumber\\
\stackrel{(*)}{=}&\frac{\nBE^+(1-\eta)+\nBE^-\eta}{\nBE^+ + \nBE^-}
\end{align}  
where in $(*)$ we used the fact that the expected value of a binomial distribution $\mathcal{B}(\nBE,\eta)$ is $\nBE \eta$.

>>>>

% Our code is available anonymously here: \url{https://anonymous.4open.science/r/GASP-4EDF}.

@@

\paragraph*{Update rules} During the agglomerative process, the interaction between adjacent clusters has to be properly updated and recomputed, as shown in Algorithm \ref{main_alg}.  
An efficient way of implementing these updates can be achieved by representing the agglomeration as a sequence of \emph{edge contractions} in the graph. Given a graph $\mathcal{G}(V,E,\cost)$ and a clustering $\Pi$, we define the associated \emph{contracted graph} $\tilde{\mathcal{G}}_\Pi(\tilde{V}, \tilde{E}, \tilde{\cost})$, such that there exists exactly one representative node $|\tilde{V} \cap S| = 1$ for every cluster $S \in \Pi$ . Edges in $\tilde{E}$ represent adjacency-relationships between clusters 
and the signed edge weights $\tilde{\cost}_e$ are given by inter-cluster interactions $\tilde{\cost}(e_{uv})=\interact_{S_u \cup S_v}$, where $S_u$ denotes the clustering including node $u$. 
For the linkage criteria tested in this article, when two clusters $S_u$ and $S_v$ are merged, the interactions between the new cluster $S_u \cup S_v$ and each of its neighbors depend only on the previous interactions involving $S_u$ and $S_v$. Thus, we can recompute these interactions by using an \emph{update rule} $f$ that does not involve any loop over the edges of the original graph $\mathcal{G}$:
\begin{align}
  \interact(S_u \cup S_v  \cup S_t) =& f\Big[ \interact(S_u  \cup  S_t), \interact(S_v  \cup  S_t) \Big] \\
  =& f(\tilde{\cost}(e_{ut}), \tilde{\cost}(e_{vt})) 
\end{align}
In Fig. \ref{fig:edge_contraction_and_contr_graph} we show an example of edge contraction and in Table \ref{tab:linkage_criteria_explicit} we list the update rules associated to the linkage criteria we introduced in Table \ref{tab:linkage-criteria}.

@@

\begin{figure}
% \vspace{-0.3cm}
\includegraphics[width=\linewidth]{figs/runtime-MWS.pdf}\\
% \vspace{-0.6cm}
\caption{Runtimes for different implementation of GASP with AbsMax linkage criterion. Runtimes are averaged over 5 runs.}
  \label{fig:runtimes_plot}
\end{figure}

\paragraph*{Efficiency of different GASP implementations with AbsMax linkage criteria} 
In Fig.~\ref{fig:runtimes_plot}, we compare the runtimes of three implementations of the AbsMax criteria: the implementation from \cite{wolf2019mutex} for pixel graphs (\emph{Pixel-grid implementation}) and for general graphs (\emph{Efficient graph implementation}) as well as the HC implementation with AbsMax linkage (\emph{Naive graph implementation}). The specialized implementations can exploit the properties of the underlying graph and are faster. But our generalization does not carry a large computational penalty and only requires a few extra seconds for partitioning graphs of a million nodes. Note that we have always used the most efficient implementation for the results reported in the paper. We will clarify this fact.

\paragraph{Median linkage} We implemented median linkage in our library from the beginning but did not report on it in the main paper for two reasons: we consider the other criteria to span the range of interesting behavior well; and it performs no better than some of the other criteria (like average linkage) which are faster to evaluate.
% Table~\ref{table:median_scores} compares average and median linkage on two clustering problems: 
% in terms of scores, the advantage of one linkage over the other seems to be application dependent, but the median linkage has higher runtime.

\begin{figure}
\includegraphics[width=\linewidth]{figs/counterexample_cropped.png}\\
\caption{GASP agglomeration with the Abs Max criterion: contracted edges are marked green. The last contraction increases the MC objective from -1 to 0.}\label{counterexample}
\end{figure}

\subsection{GASP relation to the multicut objective} \label{sec:relation_to_multicut}
For some of the linkage criteria, e.g. sum and average, GASP can be understood as a local search to the objective of the multicut optimization problem \ref{eq:MC_objective}, see \cite{levinkov2017comparative}. But this does not hold in general: the Abs Max linkage for example does not always decrease the MC objective (see counter example in Fig.~\ref{counterexample}). 
Moreover, GASP cannot be seen as a k-approximation, because it is a polynomial algorithm and \cite{chawla2006hardness} has shown that approximating the multicut objective with any constant factor is in itself NP-hard.

@@

\begin{lemma} \label{lemma:absMax_and_complete_not_positive}
If \algname{} Algorithm \ref{main_alg} with \textbf{AbsMax linkage criteria} enforces a constraint between two clusters in \emph{Phase 1}, then the interaction between the clusters will never become positive over the course of the following agglomeration steps.
\end{lemma}
\begin{proof}
During the agglomeration the interaction between two clusters can only increase in absolute value. Thus, the negative interaction $\interact{}(S_i \cup S_j)<0$ between two constrained clusters can possibly become positive over the course of next agglomeration steps only if there is at least another pair of clusters in the graph that has a positive interaction $\interact{}(S_l \cup S_t)>0$ higher in absolute value: $|\interact{}(S_l \cup S_t)| > |\interact{}(S_i \cup S_j)|$.
If such clusters $S_l, S_t$ with positive interaction exist, we note that they must also be constrained (in the opposite case, the algorithm would have already merged them before to constrain $S_i$ and $S_j$, because their priority is higher). In other words, a constrained negative interaction can become positive only if there is already another positive constrained interaction: but this can never be the case because initially all constrained interactions are negative.
\end{proof}

@@

The constrained versions of \algname{} (HCC-Avg and HCC-Single) are also not weight-shift invariant: here, the algorithm merges or constrains clusters in a given order, depending on the absolute interactions $|\interact{}(S_i \cup  S_j)|$ between clusters; so, when edge weights are shifted by a constant $\alpha$, the sorting by absolute value can change arbitrarily together with the agglomeration order, as we show in the counter-example of Fig.~\ref{fig:counter_examples_shift_inv}.
Similarly, the Mutex Watershed algorithm is not weight-shift invariant because it uses a linkage criterion that compares weights by their absolute values (see again counter-example in Fig.~\ref{fig:counter_examples_shift_inv}.
\end{proof}

@@

\algname{} puts clusters in a priority queue (Algorithm \ref{main_alg}, lines 5 and 15) and merges them starting from those with the highest interaction (lines 9, 19, and 26). However, the priority queue is updated each time two clusters are merged (lines 10, 20, and 27). Thus, to ensure a monotonously decreasing merging order, updated interactions involving a merged cluster should always be lower or equal than previously existing interactions (\textbf{condition 1}):
\begin{align}\label{eq:condition1}
& \forall S_i \in \Pi \setminus \{S_1, S_2\}, \nonumber \\
\interact{}(S_1 \cup S_2 \cup  S_i) &\leq \max \{ \interact{}(S_1  \cup  S_i), \interact{}(S_2  \cup  S_i)\} 
\end{align}
where $\Pi$ is a clustering, $\interact{}$ is a linkage criteria, and $S_1,S_2\in \Pi$ are two clusters merged by the algorithm at a given iteration. If this condition is true then, in the following iterations, \algname{} can only merge clusters with lower (or equal) interaction values.

@@

\subsection{Mutex Watershed on SSBM graphs}
\begin{prop} \label{prop:MWS_on_SSBM}
Consider a graph generated by an Erd\H os-R\'enyi signed stochastic block model (SSBM) as described in Section \ref{sec:clustering_problems}, with $N$ nodes, edges added with probability $p$, sign-flip probability $\eta<0.5$, $k$ ground-truth clusters, and edge weights Gaussian-distributed with standard deviation $\sigma$. Then, at every iteration, \algname{} with Absolute Maximum linkage (or, in other words, the Mutex Watershed algorithm) always makes a mistake with at least probability $\eta$. 
\end{prop}
\begin{proof}
Thanks to Lemma \ref{lemma:absMax_and_complete_property} we know that \algname{} with Absolute Maximum linkage returns the same clustering whether or not cannot-link-constraints are used. Thus, in the following, we prove the proposition considering the version enforcing constraints.
Let us consider a generic iteration of the algorithm, where two clusters $S_\alpha$ and $S_\beta$ have the highest priority and are popped from priority queue. Then, the MWS algorithm will either merge or constrain them depending on the fact that their interaction $\interact_{\mathrm{AbsMax}}(S_\alpha  \cup S_\beta)$ is  positive or negative (note that, with AbsMax linkage, an interaction can never be positive and constrained, as shown in Lemma \ref{lemma:absMax_and_complete_property}).
By construction of the SSBM, every edge $e\in E$ in the graph has a absolute weight distributed as $|w_e|\sim \mathcal{N}(1,\sigma^2)$. Thus, every edge $e'\in(S_{\alpha} \times S_{\beta})\cap E$ connecting the two clusters has the same probability to have the highest absolute weight, and the sign of the interaction $\interact_{\mathrm{AbsMax}}(S_\alpha  \cup  S_\beta)$ will only depend on the sign of this highest edge. Therefore, the probability that the MWS merges two clusters is simply given by the fraction of positive weighted edges connecting them.

@@

So far we have been talking about attractive and repulsive connections according to the ground truth labels. In our SSBM however every edge has a uniform probability $\eta$ to have its sign flip, so  the actual number of attractive interactions connecting the two clusters will be instead given by the sum of the true attractive interactions $\nBE^+_{\mathrm{nf}}\sim \mathcal{B}(\nBE^+,1-\eta)$ that have not been flipped, plus the true negative interactions $\nBE^-_{\mathrm{f}}\sim \mathcal{B}(\nBE^-,\eta)$ that have been flipped.
Putting everything together, given two clusters with $\nBE^+$ true attractive interactions and $\nBE^-$ true negative ones, the highest-absolute-weight edge connecting them has the following probability to be positive:
\begin{align}\label{eq:mws_prob}
 \mathbb{P}[\interact_{\mathrm{AbsMax}}&(S_\alpha  \cup  S_\beta)>0;\nBE^+,\nBE^-] = \nonumber \\
=&\sum_{\nBE^+_{\mathrm{nf}}=0}^{\nBE^+}\sum_{\nBE^-_{\mathrm{f}}=0}^{\nBE^-} 
\mathcal{B}(\nBE^-_{\mathrm{f}};\nBE^-,\eta) \mathcal{B}(\nBE^+_{\mathrm{nf}};\nBE^+,1-\eta)\cdot  \nonumber\\
& \qquad \qquad  \cdot \left(\frac{\nBE^+_{\mathrm{nf}}+\nBE^-_{\mathrm{f}}}{\nBE^+ + \nBE^-} \right) \nonumber\\
\stackrel{(*)}{=}&\frac{\nBE^+(1-\eta)+\nBE^-\eta}{\nBE^+ + \nBE^-}
\end{align}  
where in $(*)$ we used the fact that the expected value of a binomial distribution $\mathcal{B}(\nBE,\eta)$ is $\nBE \eta$.
